---
title: "Fine tuning hyperparameter"
author: "Maren Baumgärtner"
output: 
  html_notebook:
    toc: true
    toc_float: true
    code_folding: show
---

```{r, message=FALSE, echo=FALSE,warning=FALSE}
# Import libraries and dataset
library(knitr)
library(plyr)
library(dplyr)
library(reshape2)
library(randomForest)
library(ggplot2)
library(caTools)
library(mlogit)
library(nnet)
library(caret)
library(glmnet)
library(naivebayes)
library(rpart)
library(randomForest)
library(MASS)
library(FNN)
library(e1071)
library(igraph)
library(caret)
library(sparklyr)
library(RSSL)
library(sgd)
library(MASS)
library(hdm)
library(predtools)
library(xgboost)
library(adabag)
library(rpart)
library(grf)
library(ada)
library(fastAdaboost)
library(causalDML)
library(binda)
library(kknn)
library(calibrate)
library(bartMachine)
library(ranger)
```

```{r}
source("ml_wrapper.R")
source("eval_functions.R")

set.seed(42)
options(scipen= 999)
```

Define inherently multi-valued classifieres

```{r}
classifiers <- c(
  "bagging", "knn", "knn_radius", "lda", "logit",
  "mlpc", "multinom", "nb_bernulli", "nb_gaussian",
  "probability_forest", "qda", "ranger", "xgboost"
)
```

# Create multi-valued train data

```{r}
n=2000
K=3
p=15

W <- sample(seq(1, K, by = 1), replace = TRUE, n) # random W
X = matrix(runif(n*p,-pi,pi),ncol=p)
#X = matrix(sample(-0:2, n * p, replace = TRUE), nrow = n, ncol = p)

# Define a function to assign probabilities based on the sign of X[1]
assign_probabilities <- function(x) {
  p_class_1 <- ifelse(x[1] < 0, 0.7, 0.15)
  p_class_3 <- ifelse(x[1] > 0, 0.7, 0.15)
  p_class_2 <- 1 - p_class_1 - p_class_3
  probabilities <- c(p_class_1, p_class_2, p_class_3)
  return(probabilities)
}

# Apply the function to each row of X to get probabilities for W
probabilities <- t(apply(X, 1, assign_probabilities))

# Randomly sample W based on the probabilities
W <- apply(probabilities, 1, function(p) sample(1:K, size = 1, prob = p))

W.levels <- sort(unique(W))
K <- length(W.levels)

dataset <- data.frame(W=W,X)

split_indices <- sample.split(dataset$W, SplitRatio = 0.8)
training_data <- dataset[split_indices, ]
test_data <- dataset[!split_indices, ]

W_training <- as.matrix(data.frame(W = training_data$W))
X_training <- as.matrix(data.frame(X[split_indices, ]))
W_test <-     as.matrix(data.frame(W = test_data$W))
X_test <-     as.matrix(data.frame(X[!split_indices, ]))
```

# Define methods

## 1. Multinomial logistic regression (glmnet)

-   alpha: Mixing Percentage

-   lambda: Regularization

```{r}
# define new method
multinomial_logit <- list(type = "Classification",
              library = "glmnet",
              multiclass = TRUE) 
# define parameters
multinomial_logit$parameters <- data.frame(parameter = c("alpha", "lambda"),
                  class = rep("numeric", 2),
                  label = c("Mixing Percentage", "Regularization"))

multinomial_logit$fit <- logit_fit
multinomial_logit$predict <- predict.logit_fit
# design the parameter tuning grid
multinomial_logit$grid <- expand.grid(alpha = seq(0,1,0.5),
                    lambda = seq(0,1,0.5))


#test_model <- multinomial_logit$fit(X_training,W_training, alpha=0, lambda=0)
#test_pred <- multinomial_logit$predict(test_model, x=X_test, y=W_test)
```

## 2. Multinomial logistic regression (nnet)

multinom calls nnet() with zero hidden layers (size=0)

-   alpha: Mixing Percentage

-   lambda: Regularization

```{r}
# define new method
multinomial_logit_nnet <- list(type = "Classification",
              library = "nnet",
              multiclass = TRUE) 
# define parameters
multinomial_logit_nnet$parameters <- data.frame(parameter = c("alpha", "lambda"),
                  class = rep("numeric", 2),
                  label = c("Mixing Percentage", "Regularization"))

multinomial_logit_nnet$fit <- logit_fit
multinomial_logit_nnet$predict <- predict.logit_fit
# design the parameter tuning grid
multinomial_logit_nnet$grid <- expand.grid(alpha = seq(0,1,0.5),
                    lambda = seq(0,1,0.5))

```

## 3. kNN

-   kmax: number of neighbors considered

-   distance: parameter of Minkowski distance

```{r}
knn <- list(type = "Classification",
            library = "kknn",
            multiclass = TRUE)
params <- data.frame(parameter = c("kmax", "distance", "kernel"),
                     class = c("numeric", "numeric", ""),
                     label = c("number of neighbors considered", 
                               "parameter of Minkowski distance",
                               "Kernel to use"))
knn$params <- params
knn$fit <- knn_fit
knn$predict <- predict.knn_fit
knn$grid <- expand.grid(kmax = seq(5,30,10),
                        #distance = seq(1,5,1),
                        kernel = c("rectangular", "gaussian", "optimal"))

#test_model <- knn$fit(X_training,W_training, kmax=5, distance=1, kernel = "rectangular")
#test_pred <- knn$predict(test_model, x=X_test, y=W_test)
```

## 4. kNN (Radius)

```{r}
knn_radius <- list(type = "Classification",
            library = "kknn",
            multiclass = TRUE)
params <- data.frame(parameter = c("distance", "kernel"),
                     class = c("numeric", ""),
                     label = c("parameter of Minkowski distance",
                               "Kernel to use"))
knn_radius$params <- params
knn_radius$fit <- knn_fit
knn_radius$predict <- predict.knn_fit
knn_radius_grid <- expand.grid(distance = seq(1,5,1),
                        kernel = c("rectangular", "gaussian", "optimal"))
knn_radius$grid <- knn_radius_grid

#test_model <- knn_radius$fit(X_training,W_training, distance=1, kernel = "rectangular")
#test_pred <- knn_radius$predict(test_model, x=X_test, y=W_test)
```

## 5. Naive Bayes (Bernulli)

-   laplace:

```{r}
nb_bernulli <- list(
  type = "Classification",
  library = "naivebayes",
  multiclass = TRUE,
  params = data.frame(parameter = c("laplace"), 
                      class = c("numeric"), 
                      label = c("value of Laplace smoothing")),
  fit = nb_bernulli_fit,
  predict = predict.nb_bernulli_fit,
  grid = expand.grid(laplace = seq(0, 20, 2))
)
```

## 6. Naive Bayes (Gaussian)

-   laplace:

```{r}
nb_gaussian <- list(
  type = "Classification",
  library = "naivebayes",
  multiclass = TRUE,
  params = data.frame(parameter = c("laplace"), class = c("numeric"), label = c("value of Laplace smoothing")),
  fit = nb_gaussian_fit,
  predict = predict.nb_gaussian_fit,
  grid = expand.grid(laplace = seq(0, 20, 2))
)
```

## 7. Probability forest (grf)

-   mtry:

-   min.node.size:

-   alpha: controls maximum imbalance of split

-   imbalance penalty: controls how harshly imbalanced splits are penalized

```{r}
probability_forest_grf <- list(
    type = "Classification",
    library = "grf",
    multiclass = TRUE,
    params = data.frame(parameter = c("mtry", "min.node.size", "alpha"), class = c("numeric", "numeric", "numeric"), label = c("")),
    fit = probability_forest_fit,
    predict = predict.probability_forest_fit,
    grid = expand.grid(mtry = seq(as.integer(sqrt(ncol(X))), as.integer(sqrt(ncol(X)))+6,2),
    min.node.size = seq(5, 20, 5),
    alpha = seq(0.05,0.1,0.03)
    )
    )
```

## 8. Probability forest (ranger)

-   mtry:

-   min.node.size:

-   alpha: controls maximum imbalance of split

```{r}
probability_forest_ranger <- list(
  type = "Classification",
  library = "ranger",
  multiclass = TRUE,
  params = data.frame(parameter = c(""), class = c(""), label = c("")),
  fit = ranger_fit,
  predict = predict.ranger_fit,
  grid = expand.grid(mtry = seq(as.integer(sqrt(ncol(X))), as.integer(sqrt(ncol(X)))+6,2),
                     min.node.size = seq(5, 20, 5),
                     alpha = seq(0.05,0.1,0.03)
                     )
)
```

## 9. MLPC

-   size:

```{r}
mlpc <- list(
  type = "Classification",
  library = "nnet",
  multiclass = TRUE,
  params = data.frame(parameter = c("size"), class = c("numeric"), label = c("number of hidden layers")),
  fit = mlpc_fit,
  predict = predict.mlpc_fit,
  grid = expand.grid()
)
```

## 10. xgboost

-   

```{r}
xgboost <- list(
  type = "Classification",
  library = "xgboost",
  multiclass = TRUE,
  params = data.frame(parameter = c(""), class = c(""), label = c("")),
  fit = xgboost_fit,
  predict = predict.xgboost_fit,
  grid = expand.grid()
)
```

# Fine tuning classifier

To adapt in ml_wrapper: - fit: add ... in function arg - predict: just return fit

```{r}
best_logit <- finetuning_soft_classifier(x=X, y = W, method = multinomial_logit)

best_logit_nnet <- finetuning_soft_classifier(x=X, y = W, method = multinomial_logit_nnet)

best_knn <- finetuning_soft_classifier(x=X, y = W, method = knn)

best_knn_radius <- finetuning_soft_classifier(x=X, y = W, method = knn_radius)

best_nb_bernulli <- finetuning_soft_classifier(x=X, y=W, method = nb_bernulli)

best_nb_gaussian <- finetuning_soft_classifier(x=X, y=W, method = nb_gaussian)

best_probability_forest_grf <- finetuning_soft_classifier(x=X,y=W, method = probability_forest_grf)
```

```{r}
best_knn$best_params
```
